Deciding Bitvector Termination with Program Synthesis

### 2014-07-04 papers/termination/paper.tex
Rephrasing, fix bad hbox
GP stuff

\subsection{Genetic Programming and Incremental Evolution} \label{sec:gp}
To generate candidate solutions in the synthesise phase of the refinement loop,
we can use genetic programming (GP).  The general idea is to start with a population
of random programs and evolve a good one.  ``Good'' in this case means that the program
satisfies our specification.  To do this evolution, we compute a fitness for each program
and repeatedly apply genetic operators to the population based on this fitness.  The operators
used are {\sc crossover} and {\sc mutate}.  The crossover operator takes two programs
and combines their code in some what to create a new program.  The mutate operator takes a single
program and randomly changes parts of its code, again creating a new program.  We arrange
things so that the fitter a program is, the more likely it is to be selected as a parent
that will be crossover'd to produce a child in the next generation.  Our fitness function
is just the number of test cases on which the program met the specification.

>>>>

\subsection{Genetic Programming and Incremental Evolution} \label{sec:gp}
To generate candidate (non-)termination proofs in the synthesise phase of the refinement loop,
we make use of genetic programming (GP) \cite{langdon:fogp} as an alternative to model checking and exhaustive search.  
Genetic programming provides an intelligent and adaptive way of searching through the space of computer programs
for an individual that is highly fit in solving the problem at hand
by using the Darwinian principle of survival and reproduction of the fittest.
Thus, the program that emerges from the genetic programming paradigm is a consequence of fitness.

In our setting, we start with a population of random programs of size ..
We then genetically breed the population of these programs by applying the genetic operators 
{\sc crossover} and {\sc mutate}. The crossover operator takes two programs
and combines their code in some what to create a new program.  The mutate operator takes a single
program and randomly changes parts of its code, again creating a new program. 
Both these operators are applied to programs selected from the population in 
proportion to their observed fitness, i.e. a program is fit if it satisfies the (non-)termination specification.
Thus, our fitness function counts the number of test cases on which a program met the given specification.
Essentially, we plan a trajectory through the space of computer programs that will lead to programs with improved fitness.
%by considering those that meet the (non-)termination specification on more inputs until.

\todo{Talk about the fitness landscape?}

A potential issue might be that, in our setting, a fitness landscape isn't really very smooth, e.g. small changes in program representation
don't correspond to small changes in fitness. However, genetic programming's efficiency was shown to be unrelated to the smoothness 
of the fitness landscape \cite{langdon:fogp}.

\todo{mention incremental GP?}

%% The general idea is to start with a population
%% of random programs and evolve a good one.  ``Good'' in this case means that the program
%% satisfies our specification.  To do this evolution, we compute a fitness for each program
%% and repeatedly apply genetic operators to the population based on this fitness.  The operators
%% used are {\sc crossover} and {\sc mutate}.  The crossover operator takes two programs
%% and combines their code in some what to create a new program.  The mutate operator takes a single
%% program and randomly changes parts of its code, again creating a new program.  We arrange
%% things so that the fitter a program is, the more likely it is to be selected as a parent
%% that will be crossover'd to produce a child in the next generation.  Our fitness function
%% is just the number of test cases on which the program met the specification.
